---
title: "pyscenic"
---

## Pyscenic

### Run pyscenic

We calculated transcription factor activity based on spatial transcriptomics, and the specific calculation method has been described in the Method section.

```{r  eval=FALSE}
pkgs <- c("jhtools", "glue", "tidyverse", "jhuanglabRNAseq")
for (pkg in pkgs){
  suppressPackageStartupMessages(library(pkg, character.only = T))
}
project <- "embryo"
dataset <- "zhangjing"
species <- "human"
workdir <- glue("~/projects/{project}/analysis/{dataset}/{species}/rnaseq") %>% checkdir()
setwd(workdir)

tflist <- "/cluster/home/ztao_jh/projects/be_a_rich_man/RcisTarget/tflist/allTFs_hg38.txt"
glue::glue("/cluster/home/ztao_jh/projects/embryo/analysis/zhangjing/human/rnaseq/pyscenic/human") %>% checkdir()

slurmR::Slurm_lapply(c("yao1", "yao2", "yao6", "yao5"), function(sample_id){
  out_dir <- glue::glue("/cluster/home/ztao_jh/projects/embryo/analysis/zhangjing/human/rnaseq/pyscenic/human/{sample_id}") %>% checkdir()
  loom_dir <- glue::glue("/cluster/home/ztao_jh/projects/embryo/analysis/zhangjing/human/rnaseq/Spatial_counts/{sample_id}.csv.loom")
  cmd <- glue::glue("pyscenic grn --num_workers 20 --output {out_dir}/{sample_id}.adj.sc.tsv --method grnboost2 {loom_dir} {tflist}")
  system(cmd)
}, mc.cores = 4L, njobs = 60L, plan = "collect",
export = c("tflist"),
tmp_path = "/cluster/home/ztao_jh/projects/embryo/analysis/zhangjing/human/rnaseq/pyscenic/human", overwrite = T)


feature_dir <- "/cluster/home/ztao_jh/projects/be_a_rich_man/RcisTarget/human/hg38/hg38_500bp_up_100bp_down_full_tx_v10_clust.genes_vs_motifs.rankings.feather"
annotations_dir <- "/cluster/home/ztao_jh/projects/be_a_rich_man/RcisTarget/annotation/motifs-v10nr_clust-nr.hgnc-m0.001-o0.0.tbl"
slurmR::Slurm_lapply(c("yao1", "yao2", "yao6", "yao5"), function(sample_id){
  adj.sample_dir <- glue::glue("/cluster/home/ztao_jh/projects/embryo/analysis/zhangjing/human/rnaseq/pyscenic/human/{sample_id}/{sample_id}.adj.sc.tsv")
  loom_dir <- glue::glue("/cluster/home/ztao_jh/projects/embryo/analysis/zhangjing/human/rnaseq/Spatial_counts/{sample_id}.csv.loom")
  out_dir <- glue::glue("/cluster/home/ztao_jh/projects/embryo/analysis/zhangjing/human/rnaseq/pyscenic/human/{sample_id}/{sample_id}_reg_v10_v10hg38_500_100_rank.csv")

  cmd <- glue::glue("pyscenic ctx {adj.sample_dir} {feature_dir} --annotations_fname {annotations_dir} --expression_mtx_fname {loom_dir} --mode 'dask_multiprocessing' --output {out_dir} --num_workers 20 --mask_dropouts")
  system(cmd)
}, mc.cores = 4L, njobs = 60L, plan = "collect",
export = c("feature_dir", "annotations_dir"),
tmp_path = "/cluster/home/ztao_jh/projects/embryo/analysis/zhangjing/human/rnaseq/pyscenic/human", overwrite = T)
adj.sample_dir <- glue::glue("/cluster/home/ztao_jh/projects/embryo/analysis/zhangjing/human/rnaseq/pyscenic/human/{sample_id}/{sample_id}.adj.sc.tsv")
loom_dir <- glue::glue("/cluster/home/ztao_jh/projects/embryo/analysis/zhangjing/human/rnaseq/Spatial_counts/{sample_id}.csv.loom")
out_dir <- glue::glue("/cluster/home/ztao_jh/projects/embryo/analysis/zhangjing/human/rnaseq/pyscenic/human/{sample_id}/{sample_id}_reg_v10_v10hg38_500_100_rank.csv")

cmd <- glue::glue("pyscenic ctx {adj.sample_dir} {feature_dir} --annotations_fname {annotations_dir} --expression_mtx_fname {loom_dir} --mode 'dask_multiprocessing' --output {out_dir} --num_workers 20 --mask_dropouts")
system(cmd)

slurmR::Slurm_lapply(c("yao1", "yao2", "yao6", "yao5"), function(sample_id){
  loom_dir <- glue::glue("/cluster/home/ztao_jh/projects/embryo/analysis/zhangjing/human/rnaseq/Spatial_counts/{sample_id}.csv.loom")
  reg_dir <- glue::glue("/cluster/home/ztao_jh/projects/embryo/analysis/zhangjing/human/rnaseq/pyscenic/human/{sample_id}/{sample_id}_reg_v10_v10hg38_500_100_rank.csv")
  out_dir <- glue::glue("/cluster/home/ztao_jh/projects/embryo/analysis/zhangjing/human/rnaseq/pyscenic/human/{sample_id}/{sample_id}_SCENIC.csv")
  cmd <- glue::glue("pyscenic aucell {loom_dir} {reg_dir} --output {out_dir} --num_workers 8")
  system(cmd)
}, mc.cores = 4L, njobs = 60L, plan = "collect",
export = c("feature_dir", "annotations_dir"),
tmp_path = "/cluster/home/ztao_jh/projects/embryo/analysis/zhangjing/human/rnaseq/pyscenic/human", overwrite = T)


tflist <- "/cluster/home/ztao_jh/projects/be_a_rich_man/RcisTarget/tflist/allTFs_mm.txt"
glue::glue("/cluster/home/ztao_jh/projects/embryo/analysis/zhangjing/human/rnaseq/pyscenic/mouse") %>% checkdir()

slurmR::Slurm_lapply(c("ME9.5", "ME11.5x1", "ME11.5x2", "ME13.5"), function(sample_id){
  out_dir <- glue::glue("/cluster/home/ztao_jh/projects/embryo/analysis/zhangjing/human/rnaseq/pyscenic/mouse/{sample_id}") %>% checkdir()
  loom_dir <- glue::glue("/cluster/home/ztao_jh/projects/embryo/analysis/zhangjing/human/rnaseq/Spatial_counts/{sample_id}.csv.loom")
  cmd <- glue::glue("pyscenic grn --num_workers 20 --output {out_dir}/{sample_id}.adj.sc.tsv --method grnboost2 {loom_dir} {tflist}")
  system(cmd)
}, mc.cores = 4L, njobs = 60L, plan = "collect",
export = c("tflist"),
tmp_path = "/cluster/home/ztao_jh/projects/embryo/analysis/zhangjing/human/rnaseq/pyscenic/mouse", overwrite = T)


feature_dir <- "/cluster/home/ztao_jh/projects/be_a_rich_man/RcisTarget/mouse/mm10/mm10_500bp_up_100bp_down_full_tx_v10_clust.genes_vs_motifs.rankings.feather"
annotations_dir <- "/cluster/home/ztao_jh/projects/be_a_rich_man/RcisTarget/annotation/motifs-v10nr_clust-nr.mgi-m0.001-o0.0.tbl"
slurmR::Slurm_lapply(c("ME9.5", "ME11.5x1", "ME11.5x2", "ME13.5"), function(sample_id){
  adj.sample_dir <- glue::glue("/cluster/home/ztao_jh/projects/embryo/analysis/zhangjing/human/rnaseq/pyscenic/mouse/{sample_id}/{sample_id}.adj.sc.tsv")
  loom_dir <- glue::glue("/cluster/home/ztao_jh/projects/embryo/analysis/zhangjing/human/rnaseq/Spatial_counts/{sample_id}.csv.loom")
  out_dir <- glue::glue("/cluster/home/ztao_jh/projects/embryo/analysis/zhangjing/human/rnaseq/pyscenic/mouse/{sample_id}/{sample_id}_reg_v10_v10mm10_500_100_rank.csv")

  cmd <- glue::glue("pyscenic ctx {adj.sample_dir} {feature_dir} --annotations_fname {annotations_dir} --expression_mtx_fname {loom_dir} --mode 'dask_multiprocessing' --output {out_dir} --num_workers 20 --mask_dropouts")
  system(cmd)
}, mc.cores = 4L, njobs = 60L, plan = "collect",
export = c("feature_dir", "annotations_dir"),
tmp_path = "/cluster/home/ztao_jh/projects/embryo/analysis/zhangjing/human/rnaseq/pyscenic/mouse", overwrite = T)


slurmR::Slurm_lapply(c("ME9.5", "ME11.5x1", "ME11.5x2", "ME13.5"), function(sample_id){
  loom_dir <- glue::glue("/cluster/home/ztao_jh/projects/embryo/analysis/zhangjing/human/rnaseq/Spatial_counts/{sample_id}.csv.loom")
  reg_dir <- glue::glue("/cluster/home/ztao_jh/projects/embryo/analysis/zhangjing/human/rnaseq/pyscenic/mouse/{sample_id}/{sample_id}_reg_v10_v10mm10_500_100_rank.csv")
  out_dir <- glue::glue("/cluster/home/ztao_jh/projects/embryo/analysis/zhangjing/human/rnaseq/pyscenic/mouse/{sample_id}/{sample_id}_SCENIC.loom")
  cmd <- glue::glue("pyscenic aucell {loom_dir} {reg_dir} --output {out_dir} --num_workers 8")
  system(cmd)
}, mc.cores = 4L, njobs = 60L, plan = "collect",
tmp_path = "/cluster/home/ztao_jh/projects/embryo/analysis/zhangjing/human/rnaseq/pyscenic/mouse", overwrite = T)
```

### Check TF on tissue and Run TF Moran's I

To identify transcription factors associated with embryonic development, we computed the Moran's Index scores of transcription factor activity scores across different stages of embryonic development. We clustered transcription factor scores and mapped their distribution and associated modules across tissues.

```{r eval=FALSE}

pkgs <- c("jhtools", "glue", "tidyverse", "jhuanglabRNAseq","Seurat","viridis","ComplexHeatmap")
for (pkg in pkgs){
  suppressPackageStartupMessages(library(pkg, character.only = T))
}
project <- "embryo"
dataset <- "zhangjing"
species <- "human"
workdir <- glue("~/projects/{project}/analysis/{dataset}/{species}/rnaseq") %>% checkdir()
setwd(workdir)

gene_human_mouse <- read_rds("./MoransI/gene_human_mouseq.rds")
gene_ml_hu <- gene_human_mouse[["gene_ml_hu"]]
huch <- tibble(tissue_label = names(gene_ml_hu), new = c("yao1", "yao2", "yao5", "yao6"))
names(gene_ml_hu) <- huch$new
gene_ml_hu <- gene_ml_hu[c("yao1", "yao6", "yao5")]

gene_ml_mu <- gene_human_mouse[["gene_ml_mu"]]
much <- tibble(tissue_label = names(gene_ml_mu), new = c("ME9.5", "ME13.5", "ME11.5x2", "ME11.5x1"))
names(gene_ml_mu) <- much$new
gene_ml_mu <- gene_ml_mu[c("ME9.5", "ME11.5x1", "ME13.5")]

humanobj <- "/cluster/home/danyang_jh/projects/collabrators/analysis/wangwenjie/human/visium/manual_anot/seu_mrg2.rds" %>% read_rds()
DefaultAssay(humanobj) <- "Spatial"
humanobj <- JoinLayers(humanobj)
humanobj <- humanobj %>% NormalizeData() %>% ScaleData()
humanobj@meta.data$tissue_label <- paste0(humanobj$orig.ident, humanobj$stage)
rn <- rownames(humanobj@meta.data)
humanobj@meta.data <- humanobj@meta.data %>% left_join(as.data.frame(huch))
rownames(humanobj@meta.data) <- rn

mouseobj <- "/cluster/home/danyang_jh/projects/collabrators/analysis/wangwenjie/mouse/visium/integrat/mtb_genes/anot_new/seu_mrg2.rds" %>% read_rds()
delbarcode <- read_csv("delE95.csv")
mouseobj <- mouseobj[,!colnames(mouseobj) %in% delbarcode[[1]]]

DefaultAssay(mouseobj) <- "Spatial"
mouseobj <- JoinLayers(mouseobj)
mouseobj <- mouseobj %>% NormalizeData() %>% ScaleData()
mouseobj@meta.data$tissue_label <- paste0(mouseobj$orig.ident, mouseobj$stage)
rn <- rownames(mouseobj@meta.data)
mouseobj@meta.data <- mouseobj@meta.data %>% left_join(as.data.frame(much))
rownames(mouseobj@meta.data) <- rn

humanobjlist <- Seurat::SplitObject(humanobj, split.by = "new")
mouseobjlist <- Seurat::SplitObject(mouseobj, split.by = "new")

parallel::mclapply(c("ME9.5", "ME11.5x1", "ME11.5x2", "ME13.5"), function(sample_id){
  SCENIC <- glue::glue("/cluster/home/ztao_jh/projects/embryo/analysis/zhangjing/human/rnaseq/pyscenic/mouse/{sample_id}/{sample_id}_SCENIC.csv") %>% 
    read_csv()
  obj <- mouseobjlist[[sample_id]]
  SCENIC <- SCENIC %>% column_to_rownames(var = "Cell") %>% t() %>% as.data.frame()
  SCENIC <- SCENIC[,colnames(obj)]
  SCENIC <- SCENIC[rowSums(SCENIC) > 0,]
  # SCENIC <- SCENIC[,c(T, colSums(SCENIC[-1]) > 0)]
  TFl <- rownames(SCENIC)
  
  obj[["SCENIC"]] <- Seurat::CreateAssayObject(data = as.matrix(SCENIC))

  pth <- glue::glue("/cluster/home/ztao_jh/projects/embryo/analysis/zhangjing/human/rnaseq/pyscenic/mouse/{sample_id}/TF_tissue") 
  dir.create(pth, recursive = T)
  parallel::mclapply(TFl, function(i){
  p <- Seurat::SpatialFeaturePlot(obj, features = i, pt.size.factor = 3) + 
      theme_classic()+
      ggtitle(i)  +
      coord_fixed() +
      scale_fill_viridis() 
    pdf(glue("{pth}/{i}.pdf"), width = 5, height = 5)
    print(p)
    dev.off()
  }, mc.cores = 10L)
}, mc.cores = 4L)


parallel::mclapply(c("yao1", "yao2", "yao5", "yao6"), function(sample_id){
  SCENIC <- glue::glue("/cluster/home/ztao_jh/projects/embryo/analysis/zhangjing/human/rnaseq/pyscenic/human/{sample_id}/{sample_id}_SCENIC.csv") %>% 
    read_csv()
  obj <- humanobjlist[[sample_id]]
  SCENIC <- SCENIC %>% column_to_rownames(var = "Cell") %>% t() %>% as.data.frame()
  SCENIC <- SCENIC[,colnames(obj)]
  SCENIC <- SCENIC[rowSums(SCENIC) > 0,]
  TFl <- rownames(SCENIC)
  
  obj[["SCENIC"]] <- Seurat::CreateAssayObject(data = as.matrix(SCENIC))
  
  pth <- glue::glue("/cluster/home/ztao_jh/projects/embryo/analysis/zhangjing/human/rnaseq/pyscenic/human/{sample_id}/TF_tissue") 
  dir.create(pth, recursive = T)
  parallel::mclapply(TFl, function(i){
    if(sample_id %in% c("yao5", "yao6")){
      p <- Seurat::SpatialFeaturePlot(obj, features = i, pt.size.factor = 2) + 
        theme_classic()+
        ggtitle(i)  +
        coord_fixed() +
        scale_fill_viridis() 
    }else{
      p <- Seurat::SpatialFeaturePlot(obj, features = i, pt.size.factor = 3) + 
        theme_classic()+
        ggtitle(i)  +
        coord_fixed() +
        scale_fill_viridis() 
    }

    pdf(glue("{pth}/{i}.pdf"), width = 5, height = 5)
    print(p)
    dev.off()
  }, mc.cores = 10L)
}, mc.cores = 4L)


molI_frame <- parallel::mclapply(c("ME9.5", "ME11.5x1", "ME11.5x2", "ME13.5"), function(sample_id){
  SCENIC <- glue::glue("/cluster/home/ztao_jh/projects/embryo/analysis/zhangjing/human/rnaseq/pyscenic/mouse/{sample_id}/{sample_id}_SCENIC.csv") %>% 
    read_csv()
  obj <- mouseobjlist[[sample_id]]
  SCENIC <- SCENIC[,c(T, colSums(SCENIC[-1]) > 0)]
  SCENIC <- SCENIC[SCENIC$Cell %in% colnames(obj),]
  pos <- obj@images[[1]]@coordinates[,c("imagerow", "imagecol")]
  molI <- Seurat::RunMoransI(scale(SCENIC[-1]) %>% t(), pos = pos)
  molI
}, mc.cores = 4L)
names(molI_frame) <- c("ME9.5", "ME11.5x1", "ME11.5x2", "ME13.5")

molI_frame_hu <- parallel::mclapply(c("yao1", "yao2", "yao5", "yao6"), function(sample_id){
  SCENIC <- glue::glue("/cluster/home/ztao_jh/projects/embryo/analysis/zhangjing/human/rnaseq/pyscenic/human/{sample_id}/{sample_id}_SCENIC.csv") %>% 
    read_csv()
  obj <- humanobjlist[[sample_id]]
  SCENIC <- SCENIC[,c(T, colSums(SCENIC[-1]) > 0)]
  SCENIC <- SCENIC[SCENIC$Cell %in% colnames(obj),]
  pos <- obj@images[[1]]@coordinates[,c("imagerow", "imagecol")]
  molI <- Seurat::RunMoransI(SCENIC[-1] %>% scale() %>% t(), pos = pos)
  molI
}, mc.cores = 4L)
names(molI_frame_hu) <- c("yao1", "yao2", "yao5", "yao6")

write_rds(molI_frame, file = glue::glue("/cluster/home/ztao_jh/projects/embryo/analysis/zhangjing/human/rnaseq/pyscenic/mouse/mouse_MoransI.rds") )
write_rds(molI_frame_hu, file = glue::glue("/cluster/home/ztao_jh/projects/embryo/analysis/zhangjing/human/rnaseq/pyscenic/human/human_MoransI.rds") )

molI_frame_ft <- lapply(molI_frame, function(x){
  x %>% dplyr::filter(observed > 0.2) %>% rownames_to_column("TF")
})
molI_frame_hu_ft <- lapply(molI_frame_hu, function(x){
  x %>% dplyr::filter(observed > 0.2) %>% rownames_to_column("TF")
})

cluster_tf <- function(obj, spmolI, k = 7, fn){
  SCENIC <- SCENIC[,c(T, colSums(SCENIC[-1]) > 0)]
  SCENIC <- SCENIC[SCENIC$Cell %in% colnames(obj),]
  SCENIC <- SCENIC %>% column_to_rownames(var = "Cell") %>% dplyr::select(any_of(spmolI))
  
  SCENIC_scale <- SCENIC %>% scale
  corm <- cor(SCENIC_scale, method = "spearman")
  hc <- hclust(as.dist(1-corm), method = "ward.D2")
  
  htd <- Heatmap(corm,
                 cluster_columns = hc, 
                 cluster_rows = hc, 
                 column_split = k,
                 row_split = k)
  pdf(fn, width = 30, height = 30)
  print(htd)
  dev.off()

  odc <- column_order(htd)
  dev.off()
  lapply(1:length(odc), function(i){
    tibble(
      tf = colnames(corm)[odc[[i]]],
      group = paste0("cluster_", i)
    )
  }) %>% 
    list_rbind()
}
aggrated_value <- function(SCENIC, cluster_tf_df, obj){
  summarise_tf_cluster <- SCENIC %>% 
    pivot_longer(cols = -Cell, names_to = "tf") %>% 
    left_join(cluster_tf_df) %>% 
    na.omit() %>% 
    group_by(group, Cell) %>% 
    summarise(value = mean(value)) %>% 
    ungroup()
  img_summarise_tf_cluster <- summarise_tf_cluster %>%
    left_join(
      obj@images[[1]]@coordinates[,c("imagerow", "imagecol")] %>% rownames_to_column(var = "Cell")
    )
  img_summarise_tf_cluster
}
plot_sp <- function(img_summarise_tf_cluster){
  p <- img_summarise_tf_cluster %>% 
    ggplot(aes(x = imagecol, y = imagerow, color = value)) +
    geom_point()  +
    theme_classic()+
    scale_color_viridis() +
    facet_wrap(~group, ncol = 4) +
    coord_fixed()
  p
}

c("ME9.5", "ME11.5x1", "ME11.5x2", "ME13.5")
sample_id <- "ME9.5"
SCENIC <- glue::glue("/cluster/home/ztao_jh/projects/embryo/analysis/zhangjing/human/rnaseq/pyscenic/mouse/{sample_id}/{sample_id}_SCENIC.csv") %>% 
  read_csv()
obj <- mouseobjlist[[sample_id]]
spmolI <- molI_frame_ft[[sample_id]]$TF
cluster_tf_df <- cluster_tf(obj, spmolI, k = 7, fn = glue::glue("/cluster/home/ztao_jh/projects/embryo/analysis/zhangjing/human/rnaseq/pyscenic/mouse/{sample_id}/cluster_heatmap.pdf"))
write_csv(cluster_tf_df, file = glue::glue("/cluster/home/ztao_jh/projects/embryo/analysis/zhangjing/human/rnaseq/pyscenic/mouse/{sample_id}/cluster_tf.csv"))
img_summarise_tf_cluster <- aggrated_value(SCENIC, cluster_tf_df, obj)
p <- plot_sp(img_summarise_tf_cluster)

pdf(glue::glue("/cluster/home/ztao_jh/projects/embryo/analysis/zhangjing/human/rnaseq/pyscenic/mouse/{sample_id}/cluster_tf.pdf"),
    width = 10, height = 5)
print(p)
dev.off()

sample_id <- "ME11.5x1"
SCENIC <- glue::glue("/cluster/home/ztao_jh/projects/embryo/analysis/zhangjing/human/rnaseq/pyscenic/mouse/{sample_id}/{sample_id}_SCENIC.csv") %>% 
  read_csv()
obj <- mouseobjlist[[sample_id]]
spmolI <- molI_frame_ft[[sample_id]]$TF
cluster_tf_df <- cluster_tf(obj, spmolI, k = 7, fn = glue::glue("/cluster/home/ztao_jh/projects/embryo/analysis/zhangjing/human/rnaseq/pyscenic/mouse/{sample_id}/cluster_heatmap.pdf"))
write_csv(cluster_tf_df, file = glue::glue("/cluster/home/ztao_jh/projects/embryo/analysis/zhangjing/human/rnaseq/pyscenic/mouse/{sample_id}/cluster_tf.csv"))
img_summarise_tf_cluster <- aggrated_value(SCENIC, cluster_tf_df, obj)
p <- plot_sp(img_summarise_tf_cluster)

pdf(glue::glue("/cluster/home/ztao_jh/projects/embryo/analysis/zhangjing/human/rnaseq/pyscenic/mouse/{sample_id}/cluster_tf.pdf"),
    width = 15, height = 10)
print(p)
dev.off()

sample_id <- "ME11.5x2"
SCENIC <- glue::glue("/cluster/home/ztao_jh/projects/embryo/analysis/zhangjing/human/rnaseq/pyscenic/mouse/{sample_id}/{sample_id}_SCENIC.csv") %>% 
  read_csv()
obj <- mouseobjlist[[sample_id]]
spmolI <- molI_frame_ft[[sample_id]]$TF
cluster_tf_df <- cluster_tf(obj, spmolI, k = 7, fn = glue::glue("/cluster/home/ztao_jh/projects/embryo/analysis/zhangjing/human/rnaseq/pyscenic/mouse/{sample_id}/cluster_heatmap.pdf"))
write_csv(cluster_tf_df, file = glue::glue("/cluster/home/ztao_jh/projects/embryo/analysis/zhangjing/human/rnaseq/pyscenic/mouse/{sample_id}/cluster_tf.csv"))
img_summarise_tf_cluster <- aggrated_value(SCENIC, cluster_tf_df, obj)
p <- plot_sp(img_summarise_tf_cluster)

pdf(glue::glue("/cluster/home/ztao_jh/projects/embryo/analysis/zhangjing/human/rnaseq/pyscenic/mouse/{sample_id}/cluster_tf.pdf"),
    width = 15, height = 10)
print(p)
dev.off()

sample_id <- "ME13.5"
SCENIC <- glue::glue("/cluster/home/ztao_jh/projects/embryo/analysis/zhangjing/human/rnaseq/pyscenic/mouse/{sample_id}/{sample_id}_SCENIC.csv") %>% 
  read_csv()
obj <- mouseobjlist[[sample_id]]
spmolI <- molI_frame_ft[[sample_id]]$TF
cluster_tf_df <- cluster_tf(obj, spmolI, k = 7, fn = glue::glue("/cluster/home/ztao_jh/projects/embryo/analysis/zhangjing/human/rnaseq/pyscenic/mouse/{sample_id}/cluster_heatmap.pdf"))
write_csv(cluster_tf_df, file = glue::glue("/cluster/home/ztao_jh/projects/embryo/analysis/zhangjing/human/rnaseq/pyscenic/mouse/{sample_id}/cluster_tf.csv"))
img_summarise_tf_cluster <- aggrated_value(SCENIC, cluster_tf_df, obj)
p <- plot_sp(img_summarise_tf_cluster)

pdf(glue::glue("/cluster/home/ztao_jh/projects/embryo/analysis/zhangjing/human/rnaseq/pyscenic/mouse/{sample_id}/cluster_tf.pdf"),
    width = 20, height = 18)
print(p)
dev.off()


c("yao1", "yao2", "yao5", "yao6")
sample_id <- "yao1"
SCENIC <- glue::glue("/cluster/home/ztao_jh/projects/embryo/analysis/zhangjing/human/rnaseq/pyscenic/human/{sample_id}/{sample_id}_SCENIC.csv") %>% 
  read_csv()
obj <- humanobjlist[[sample_id]]
spmolI <- molI_frame_hu_ft[[sample_id]]$TF
cluster_tf_df <- cluster_tf(obj, spmolI, k = 7, fn = glue::glue("/cluster/home/ztao_jh/projects/embryo/analysis/zhangjing/human/rnaseq/pyscenic/human/{sample_id}/cluster_heatmap.pdf"))
write_csv(cluster_tf_df, file = glue::glue("/cluster/home/ztao_jh/projects/embryo/analysis/zhangjing/human/rnaseq/pyscenic/human/{sample_id}/cluster_tf.csv"))
img_summarise_tf_cluster <- aggrated_value(SCENIC, cluster_tf_df, obj)
p <- plot_sp(img_summarise_tf_cluster)

pdf(glue::glue("/cluster/home/ztao_jh/projects/embryo/analysis/zhangjing/human/rnaseq/pyscenic/human/{sample_id}/cluster_tf.pdf"),
    width = 14, height = 7)
print(p)
dev.off()

sample_id <- "yao2"
SCENIC <- glue::glue("/cluster/home/ztao_jh/projects/embryo/analysis/zhangjing/human/rnaseq/pyscenic/human/{sample_id}/{sample_id}_SCENIC.csv") %>% 
  read_csv()
obj <- humanobjlist[[sample_id]]
spmolI <- molI_frame_hu_ft[[sample_id]]$TF
cluster_tf_df <- cluster_tf(obj, spmolI, k = 7, fn = glue::glue("/cluster/home/ztao_jh/projects/embryo/analysis/zhangjing/human/rnaseq/pyscenic/human/{sample_id}/cluster_heatmap.pdf"))
write_csv(cluster_tf_df, file = glue::glue("/cluster/home/ztao_jh/projects/embryo/analysis/zhangjing/human/rnaseq/pyscenic/human/{sample_id}/cluster_tf.csv"))
img_summarise_tf_cluster <- aggrated_value(SCENIC, cluster_tf_df, obj)
p <- plot_sp(img_summarise_tf_cluster)

pdf(glue::glue("/cluster/home/ztao_jh/projects/embryo/analysis/zhangjing/human/rnaseq/pyscenic/human/{sample_id}/cluster_tf.pdf"),
    width = 14, height = 10)
print(p)
dev.off()


sample_id <- "yao6"
SCENIC <- glue::glue("/cluster/home/ztao_jh/projects/embryo/analysis/zhangjing/human/rnaseq/pyscenic/human/{sample_id}/{sample_id}_SCENIC.csv") %>% 
  read_csv()
obj <- humanobjlist[[sample_id]]
spmolI <- molI_frame_hu_ft[[sample_id]]$TF
cluster_tf_df <- cluster_tf(obj, spmolI, k = 7, fn = glue::glue("/cluster/home/ztao_jh/projects/embryo/analysis/zhangjing/human/rnaseq/pyscenic/human/{sample_id}/cluster_heatmap.pdf"))
write_csv(cluster_tf_df, file = glue::glue("/cluster/home/ztao_jh/projects/embryo/analysis/zhangjing/human/rnaseq/pyscenic/human/{sample_id}/cluster_tf.csv"))
img_summarise_tf_cluster <- aggrated_value(SCENIC, cluster_tf_df, obj)
p <- plot_sp(img_summarise_tf_cluster)

pdf(glue::glue("/cluster/home/ztao_jh/projects/embryo/analysis/zhangjing/human/rnaseq/pyscenic/human/{sample_id}/cluster_tf.pdf"),
    width = 20, height = 12)
print(p)
dev.off()


sample_id <- "yao5"
SCENIC <- glue::glue("/cluster/home/ztao_jh/projects/embryo/analysis/zhangjing/human/rnaseq/pyscenic/human/{sample_id}/{sample_id}_SCENIC.csv") %>% 
  read_csv()
obj <- humanobjlist[[sample_id]]
spmolI <- molI_frame_hu_ft[[sample_id]]$TF
cluster_tf_df <- cluster_tf(obj, spmolI, k = 7, fn = glue::glue("/cluster/home/ztao_jh/projects/embryo/analysis/zhangjing/human/rnaseq/pyscenic/human/{sample_id}/cluster_heatmap.pdf"))
write_csv(cluster_tf_df, file = glue::glue("/cluster/home/ztao_jh/projects/embryo/analysis/zhangjing/human/rnaseq/pyscenic/human/{sample_id}/cluster_tf.csv"))
img_summarise_tf_cluster <- aggrated_value(SCENIC, cluster_tf_df, obj)
p <- plot_sp(img_summarise_tf_cluster)

pdf(glue::glue("/cluster/home/ztao_jh/projects/embryo/analysis/zhangjing/human/rnaseq/pyscenic/human/{sample_id}/cluster_tf.pdf"),
    width = 30, height = 17)
print(p)
dev.off()
```
